---
title: "Understanding the structure of an Econometrics study"
author: "FLORENCE"
date: "2025-01-30"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```


# **Part A: Structure of an Econometrics Study**

## **Microeconomics Research Question**
**How does an individual's income and credit history affect their probability of credit card approval?**

### **General Structure**
- **Type of Research:** Secondary
- **Variables:**
  - **Dependent:** `Acc` (Credit Card Application Accepted)
  - **Independent:** `MDR`, `Age`, `Income`, `Avgexp`, `Ownrent`, `Selfempl`
- **Econometric Model:** Logistic Regression

\[
P(Acc = 1) = \frac{e^{\beta_0 + \beta_1 MDR + \beta_2 Age + \beta_3 Income + \beta_4 Avgexp + \beta_5 Ownrent + \beta_6 Selfempl}}{1 + e^{\beta_0 + \beta_1 MDR + \beta_2 Age + \beta_3 Income + \beta_4 Avgexp + \beta_5 Ownrent + \beta_6 Selfempl}}
\]

## **Macroeconomics Research Question**
**How does economic growth affect overall credit card approval rates in an economy?**

### **General Structure**
- **Type of Research:** Secondary
- **Variables:**
  - **Dependent:** `Credit Card Approval Rate`
  - **Independent:** `GDP Growth`, `Unemployment Rate`, `Inflation Rate`, `Interest Rate`
- **Econometric Model:** Multiple Linear Regression

\[
CreditCardApprovalRate = \beta_0 + \beta_1 GDPGrowth + \beta_2 Unemployment + \beta_3 Inflation + \beta_4 InterestRate + \epsilon
\]

---

# **Part B: Modelling Econometric Data in R**
## **Loading Required Libraries and Dataset**

#### Loading Required Libraries

```{r, warning=FALSE}
library(tidyverse)  
library(dplyr)      
library(ggplot2)    
library(caret)      
library(stargazer) 
library(MatchIt)    
library(broom)      
library(glmnet)     
library(car)        
library(ipw)       
```

### Dataset

```{r}
df <- read.csv("dataset_for_assignment(2)(1).csv")
```

#### Checking for missing values and summary stats

```{r}
colSums(is.na(df))
```

```{r}
summary(df)
```


#### Checking correlations to avoid multicollinearity

```{r}
correlation_matrix <- cor(df  |>
                            select(-Acc))
print(correlation_matrix)
```


#### Removing highly correlated variables - Income or Avgexp 


#### Normalizing Income and Avgexp
```{r}
df$Income <- df$Income / max(df$Income)
df$Avgexp <- df$Avgexp / max(df$Avgexp)
```


#### Checking for Variance Inflation Factor (VIF) for multicollinearity

```{r}
vif_check <- vif(lm(Acc ~ MDR + Age + Income + Avgexp + Ownrent, data = df))
print(vif_check)
```


None of the VIF > 5 for any variable. 

#### Fitting Regularized Logistic Regression (Lasso for feature selection)

```{r}
x <- as.matrix(df  |> 
                 select(MDR, Age, Income, Avgexp, Ownrent))
y <- df$Acc
cv_model <- cv.glmnet(x, y, family = "binomial", alpha = 1) 
best_lambda <- cv_model$lambda.min
logit_model <- glmnet(x, y, family = "binomial", alpha = 1, lambda = best_lambda)
```


#### Model Summary
```{r}
print(logit_model)
```


#### Average Treatment Effect (ATE) for Income
```{r}
median_income <- median(df$Income)
df$treatment <- ifelse(df$Income > median_income, 1, 0)
```


#### Propensity Score Matching
```{r}
match_model <- matchit(treatment ~ MDR + Age + Avgexp + Ownrent, data = df, method = "nearest")
matched_data <- match.data(match_model)
ATE <- mean(matched_data$Acc[matched_data$treatment == 1]) - mean(matched_data$Acc[matched_data$treatment == 0])
print(paste("ATE after Matching:", ATE))
```


#### Inverse Probability Weighting (IPW)

```{r}
df$treatment <- as.integer(df$treatment)  
weight_model <- ipwpoint(exposure = treatment, 
                         family = "binomial",
                         link = "logit",
                         denominator = ~ MDR + Age + Avgexp + Ownrent, 
                         data = df)

```



#### Computing weighted ATE

```{r}
# Extract the computed IPW weights
df$ipw <- weight_model$ipw.weights

# Check if weights have extreme values (important for stability)
summary(df$ipw)
```

```{r}
# Compute weighted ATE
weighted_ATE <- mean(df$Acc * df$ipw, na.rm = TRUE)

# Print final ATE result
print(paste("ATE using IPW:", weighted_ATE))

```



# Logistic Regression Coefficients Plot

```{r}
# Fit logistic regression using glm() to get standard errors
logit_model_glm <- glm(Acc ~ MDR + Age + Income + Avgexp + Ownrent, data = df, family = binomial)

# Convert model summary to a tidy format
logit_coeffs <- broom::tidy(logit_model_glm)

# Plot with standard errors
ggplot(logit_coeffs, aes(x = term, y = estimate)) +
  geom_col(fill = "steelblue") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(title = "Logistic Regression Coefficients", x = "Variables", y = "Coefficient Value") +
  theme_minimal()
```

#### Approval Rate by Income Group Plot

```{r}
df |> 
  group_by(treatment) |> 
  summarise(approval_rate = mean(Acc)) |> 
  ggplot(aes(x = factor(treatment), y = approval_rate, fill = factor(treatment))) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = approval_rate - sd(approval_rate), ymax = approval_rate + sd(approval_rate)), width = 0.2) +
  labs(title = "Approval Rate by Income Group", x = "Income Group (0 = Low, 1 = High)", y = "Approval Rate") +
  scale_fill_manual(values = c("red", "blue")) +
  theme_minimal()
```

